---
title: "LandscapeSurfaceGeneration"
author: "Annabelle"
date: "5/7/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## --- Load libraries  -------------------------------------------------------------- 

```{r}
library(landscapeR)
library(tidyverse)
library(raster)           #for raster covariate data; version 2.6-7 used
library(NLMR)
library(gstat)            #for variograms/kriging; version 1.1-6 used
library(sp)
library(landscapemetrics)
## library(landscapetools)
library(Voss)
library(geoR)             #for variograms/kriging; had to fix with pain in the arse... 
library(gstat)            #for variograms/kriging; version 1.1-6 used
library(RandomFields)     #for simulating spatial random fields/spatial dependence; version 3.1.50 used

library(terra)

library(landscapetools)
library(dplyr)
library(purrr)
library(tibble)


library(waveslim)         #for multi-scale dependence; version 1.7.5 used
library(fields)           #for multi-scale dependence; version 9.6 used
library(reshape2)         #for re-formatting data; version 1.4.3 used
library(vegan)            #for multi-scale dependence; version 2.5-2 used
library(adespatial)       #for multi-scale dependence; version 0.3-0 used

```


### Algorithm that can generate surfaces of vegetation 




6000 acre area with different landcover 
landcover map then overlay landuse 


```{r}
nlm_fbm() # Simulates neutral landscapes using fractional Brownian motion (fBm). fBm is an extension of Brownian motion in which the amount of spatial autocorrelation between steps is controlled by the Hurst coefficient H

nlm_gaussianfield() # Simulates a spatially correlated random fields (Gaussian random fields) model, where one can control the distance and magnitude of spatial autocorrelation

nlm_neigh() # Simulates a neutral landscape model with land cover classes and clustering based on neighbourhood characteristics. The cluster are based on the surrounding cells. If there is a neighbouring cell of the current value/type, the target cell will more likely turned into a cell of that type/value

nlm_mpd() # Simulates a midpoint displacement neutral landscape model where the parameter roughness controls the level of spatial autocorrelation **

```


## AnnabelleLand
Summary: Create landscape with specified levels of clumpyness and then break that into groups based on numeric values
```{r}
## --- Parameters -------------------------------------------------------------- 
## Land cover name: Ag_graze, Ag_row, Timber_Hard, Timber_pine, Water, Developed 

# breakdown:
#   - grassland (50%)
#   - forest (30%)
#   - wetland (20%)

set.seed(20220105 + 4)
h4 <- nlm_gaussianfield(ncol = 300, nrow = 300, 
                     resolution = 1, 
                     nug = 1,
                     autocorr_range = 10)

germ2lc <- data.frame(value = sort(unique(values(h4)))) %>% 
  dplyr::mutate(landcover = case_when(
    value < 0.5 ~ 1,
    value < 0.8 ~ 2,
    TRUE ~ 3
  ))

lc_vals <- data.frame(value = values(h4)) %>% 
  left_join(germ2lc)

lc <- setValues(h1, lc_vals$landcover)

# Check
table(values(lc))/(300*300)


### landscapetools::show_landscape(lc)
```




## Narnia 

Controlling some parts of autocorrelation (nugget, magnitude of variation, range) and then proportion of land cover types 
Note, mutiple versions only currently seem to work if have list of proportions, which gets a bit unrully 
also need to figure out how to control proportional amount and link to class names
```{r}

## --- Demo from online -------------------------------------------------------------- 

# simulation function that has the parameters we want to alter as input
simulate_landscape = function(roughness, weighting){
    nlm_mpd(ncol = 50,
            nrow = 50,
            roughness = roughness,
            rescale = TRUE) %>%
        util_classify(weighting = weighting)
}

# paramter combinations we are interested in
param_df = expand.grid(roughness = c(0.2, 0.9),
                       weighting = list(c(0.2, 0.8), c(0.2, 0.3, 0.5))) %>%
    as_tibble()

# map over the nested tibble and use each row as input for our simulation function
nlm_list = param_df %>% pmap(simulate_landscape)

# look at the results
show_landscape(nlm_list)
plot(nlm_list)
landscapetools::show_landscape(stack(nlm_list))



## --- simple version of problem -------------------------------------------------------------- 

# simulation function that has the parameters we want to alter as input
simulate_landscape = function(nug, autocorr_range, mag_var, weighting){
    nlm_gaussianfield(
            ncol = 50,
            nrow = 50,
            nug = nug,
            autocorr_range = autocorr_range, 
            mag_var = mag_var,
            rescale = TRUE) %>%
        util_classify(n = 3, weighting = weighting, level_names = c("A", "B", "C"))
}


# paramter combinations we are interested in
param_df = expand.grid(nug = 1,
                       autocorr_range = 10, 
                       mag_var = 8, 
                       weighting = list(c(0.1, .1, 0.8), c(0.2, 0.3, 0.5), c(0.4, 0.4, 0.2))) %>% ## different amounts of the habitat
    as_tibble()

# map over the nested tibble and use each row as input for our simulation function
nlm_list = param_df %>% pmap(simulate_landscape)

# look at the results
show_landscape(nlm_list)

landscapetools::show_landscape(stack(nlm_list))





## --- final version of problem -------------------------------------------------------------- 





# simulation function that has the parameters we want to alter as input
simulate_landscape = function(nug, autocorr_range, mag_var, weighting){
    nlm_gaussianfield(
            ncol = 50,
            nrow = 50,
            nug = nug,
            autocorr_range = autocorr_range, 
            mag_var = mag_var,
            rescale = TRUE) %>%
        util_classify(n = 6, weighting = weighting, level_names = c("Ag_graze", "Ag_row", "Timber_Hard", "Timber_pine", "Water", "Developed"))
}


# paramter combinations we are interested in
param_df = expand.grid(nug = c(1,2),
                       autocorr_range = c(5,20), 
                       mag_var = c(2,4), 
   #                    weighting = list(c(0.5, .5, .1, .07, .03, 0.7), c(0.1, .1, .15, .15, .25, .25), c(00.16, 0.16, 0.16, 0.16, 0.16, .2))) %>% ## different proportions of the habitat
                        weighting = c(00.16, 0.16, 0.16, 0.16, 0.16, .2)) %>% 
    as_tibble()

# map over the nested tibble and use each row as input for our simulation function
nlm_list = param_df %>% pmap(simulate_landscape)

# look at the results
landscapetools::show_landscape(stack(nlm_list))


#run for one example... 

cair_paravel <- simulate_landscape(nug = 4, autocorr_range = 60, mag_var = 30, weighting = c(00.16, 0.16, 0.16, 0.16, 0.16, .2))

landscapetools::show_landscape(stack(cair_paravel))

cair_paravel
## Values: 1 (Ag_graze) : to 6 (Developed) 

plot(cair_paravel)

```



## Mordor
Fractal landscapes 
```{r}

## use nlm_fbm function..

# simulation function that has the parameters we want to alter as input
simulate_landscape = function(fract_dim, weighting){
    nlm_fbm(
            ncol = 50,
            nrow = 50,
            fract_dim = fract_dim,
            rescale = TRUE) %>%
        util_classify(n = 6, weighting = weighting, level_names = c("Ag_graze", "Ag_row", "Timber_Hard", "Timber_pine", "Water", "Developed"))
}


# paramter combinations we are interested in
param_df = expand.grid(fract_dim = c(.2, .5, 1, 1.5),
                       weighting = list(c(0.5, .5, .1, .07, .03, 0.7), c(0.1, .1, .15, .15, .25, .25), c(00.16, 0.16, 0.16, 0.16, 0.16, .2))) %>% ## different proportions of the habitat
   #                   weighting = c(00.16, 0.16, 0.16, 0.16, 0.16, .2)) %>% 
    as_tibble()

# map over the nested tibble and use each row as input for our simulation function
nlm_list = param_df %>% pmap(simulate_landscape)

# look at the results
landscapetools::show_landscape(stack(nlm_list))


#run for one example... 

cair_paravel <- simulate_landscape(nug = 4, autocorr_range = 60, mag_var = 30, weighting = c(00.16, 0.16, 0.16, 0.16, 0.16, .2))

landscapetools::show_landscape(stack(cair_paravel))

cair_paravel
## Values: 1 (Ag_graze) : to 6 (Developed) 

plot(cair_paravel)

```


## Westworld 
random cluster algorithms 
```{r}

### !!!!! This function isn't working... 

# simulation function that has the parameters we want to alter as input
simulate_landscape = function(p_neigh, categories, weighting){
    nlm_neigh(
            ncol = 50,
            nrow = 50,
            p_neigh = p_neigh,
            p_empty = p_empty, 
            categories = categories, 
            rescale = TRUE) %>%
        util_classify(n = 6, weighting = weighting, level_names = c("Ag_graze", "Ag_row", "Timber_Hard", "Timber_pine", "Water", "Developed"))
}


TheMaze <- simulate_landscape(p_neigh = .7, p_empty = .1, categories = 6,  weighting = c(00.16, 0.16, 0.16, 0.16, 0.16, .2))


neigh_raster <- nlm_neigh(ncol = 50, nrow = 50, p_neigh = 0.7, p_empty = 0.1,
                    categories = 5, neighbourhood = 4)

```



Readings on spatial data - https://geocompr.robinlovelace.net/spatial-class.html

```{r}
elev = rast(nrows = 6, ncols = 6, resolution = 0.5, 
            xmin = -1.5, xmax = 1.5, ymin = -1.5, ymax = 1.5,
            vals = 1:36)


grain_order = c("clay", "silt", "sand")
grain_char = sample(grain_order, 36, replace = TRUE)
grain_fact = factor(grain_char, levels = grain_order)
grain = rast(nrows = 6, ncols = 6, resolution = 0.5, 
             xmin = -1.5, xmax = 1.5, ymin = -1.5, ymax = 1.5,
             vals = grain_fact)

levels(grain)[[1]] = c(levels(grain)[[1]], wetness = c("wet", "moist", "dry"))
levels(grain)
cats(grain)

hist(elev)

```







### DiseaseLand 

https://github.com/morgankain/Land-Use_Disease_Model/blob/ce272c8e0159dc307e01a9c60421e65a06bd0e8a/sim_landscapes.R

```{r}

####################################################################
## Obtain simulated landscapes across a gradient of heterogeneity ##
####################################################################

## Thee goal here is to simulate two landscapes (one for urban area and one for tree cover) with different levels of fragmentation (varying degrees of spatial auto-correlation) 
## such that on each landscape there is the same amount of overlap between tree cover and urban area (same amount of negative correlation). 
## Couldn't find a way to do this automatically with any canned package, so this script combines the simulation of individual landscapes with a brute force rejection of configurations 
## that fail certain checks (e.g., not an appropriate value of negative correlation)

## Getting landscapes in this way is a very slow process so this script is set up to allow for saving and reloading of previously simulated landscapes 

## Just a bit of setup
landscape_predictors <- c("human_population", "tree_cover")
num_land             <- length(landscape_predictors)
which_tree           <- grep("tree", landscape_predictors)

## Landscape matrix size that is reasonable for computation
matx      <- 60
maty      <- 60

## values of spatial heterogeneity
het       <- seq(0, 0.92, by = 0.02)

## range of correlation between the two landscape attributes that will be accepted for these landscapes 
## Note: the stronger the negative correlation and the narrower the window the longer the simulation will take. 
## Settled on this range as an acceptable non-overlap that would actually succeed in finding landscapes 
corrange    <- c(-0.55, -0.45)
## For each landscape also force the average value on the landscape to be near 0.5. Scaling for human and other host densities happens from the simulated 0-1 values
## that have an average across the landscape of close to 0.5 
meanrange   <- c(0.48, 0.52)

## Include parameter values in the name of the saved file otherwise this could get confusing in the future
landname <- paste(
  "saved_input/"
  , paste(
    paste("cor",  paste(corrange, collapse = "_"), sep = "_")
    , paste("mean", paste(meanrange, collapse = "_"), sep = "_")
    #, paste("mean_1", paste(meanrange_1, collapse = "_"), sep = "_")
    #, paste("mean_2", paste(meanrange_2, collapse = "_"), sep = "_")
    , paste("dim",  paste(matx, collapse = "_"), sep = "_")
    , sep = "_")
  , "_sim_land.Rds"
  , sep = "")

if (file.exists(landname)) {
  sim_land <- readRDS(landname)
} else {
  sim_land  <- vector("list", nsim) 
  ## Need these to simulate new landscapes, but if just using previously simulated landscapes don't need load them
  library(NLMR); library(raster) # library(landscapetools)
}

for (i in seq_along(het)) {
  
  ## load a previously successful simulated landscape at this 
  if (is.null(sim_land[[i]])) {
    
    ## just filling for now with data to get large cor
    land1 <- nlm_mpd(ncol = matx, nrow = maty, roughness = 0.01, rand_dev = 10)
    land2 <- nlm_mpd(ncol = matx, nrow = maty, roughness = 0.01, rand_dev = 10)
    
    ## Calc correlation between the landscapes
    tempcor   <- cor(c(land1@data@values), c(land2@data@values))
    ## Aiming for the sweetspot of corrange. When inbetween this will turn up FALSE
    testcor   <- sum(((tempcor < corrange[1]) | (tempcor > corrange[2])))
    
    ## Also check the mean of the pixel values for each landscape. Want these to be near 0.5 so that the values span the full range
    tempmeans <- matrix(data = c(mean(c(land1@data@values)), mean(c(land2@data@values))), nrow = 1)
    ## Aiming for the sweetspot of meanrange. When both are inbetween this will turn up FALSE, FALSE
    testmeans <- sum(apply(tempmeans, 2, FUN = function (x) ((x < meanrange[1]) | (x > meanrange[2])))) 
    #testmeans <- ((tempmeans[1,1] < meanrange_1[1]) | (tempmeans[1,1] > meanrange_1[2])) +
    #  ((tempmeans[1,2] < meanrange_2[1]) | (tempmeans[1,2] > meanrange_2[2]))
    
    ## simulate landscapes until testcor and testmeans is satisfied
    while (sum(testcor, testmeans) != 0) {
      
      land1 <- nlm_mpd(ncol = matx, nrow = maty, roughness = het[i], rand_dev = 10, rescale = TRUE)
      land2 <- nlm_mpd(ncol = matx, nrow = maty, roughness = het[i], rand_dev = 10, rescale = TRUE)
      
      ## checks for these simulated landscapes
      tempcor   <- cor(c(land1@data@values), c(land2@data@values)); tempcor
      testcor   <- sum(((tempcor < corrange[1]) | (tempcor > corrange[2])))
      tempmeans <- matrix(data = c(mean(c(land1@data@values)), mean(c(land2@data@values))), nrow = 1)
      testmeans <- sum(apply(tempmeans, 2, FUN = function (x) ((x < meanrange[1]) | (x > meanrange[2]))))
      #testmeans <- ((tempmeans[1,1] < meanrange_1[1]) | (tempmeans[1,1] > meanrange_1[2])) +
      #  ((tempmeans[1,2] < meanrange_2[1]) | (tempmeans[1,2] > meanrange_2[2]))
      
    }
    
    ## extract just the landscape values
    land1 <- matrix(data = land1@data@values, nrow = matx - 1, ncol = matx - 1)
    land2 <- matrix(data = land2@data@values, nrow = matx - 1, ncol = matx - 1)
    
    ## store in the list as a successful landscape combo
    temp_array      <- array(data = 0, dim = c(matx - 1, maty - 1, 2))
    temp_array[,,1] <- land1
    temp_array[,,2] <- land2
    dimnames(temp_array)[[3]] <- landscape_predictors
    
    sim_land[[i]]        <- temp_array
    
    ## And rewrite the landscape so that this landscape doesn't need to be simulated again
    saveRDS(sim_land, landname)
    
  }
  
  ## make sure i = i + 1 fulfills while loop requirements
  testcor <- 1
  #print(het[i])
  
}

## Use the sim landscape and fill out the rest of this landsape entry
landscape[[landscape_counter]] <- array(
  data = 0
  , dim = c(
    matx - 1
    , maty - 1
    , 2 + num_hosts + num_mosq
  ))

if (disease[disease_counter] != "yellow_fever") {
  dimnames(landscape[[landscape_counter]]) <- list(
    NULL
    , NULL
    , c("urban", "human", "forest", "other", mosq_names)
  )
} else {
  dimnames(landscape[[landscape_counter]]) <- list(
    NULL
    , NULL
    , c("urban", "human", "forest", "primate", "other", mosq_names)
  )  
}

## A few final needed scaling steps

## Just the current one
current_landscape <- sim_land[[which_landscapes[landscape_counter]]]

## First, downstream breaks if there are true 0 trees or true 0 humans. I should fix
## downstream how mosquito distributions are calculated and allow 0s, but for now it breaks
## (see mosq_coverage calculation in R0_internal_calc_fast_opaque.R)
zero_vals.h <- which(current_landscape[,,landscape_predictors[1]] == 0, arr.ind = T)
nz_min.h    <- min(current_landscape[,,landscape_predictors[1]][current_landscape[,,landscape_predictors[1]] != 0])
for (zf in 1:nrow(zero_vals.h)) {
  current_landscape[,,landscape_predictors[1]][zero_vals.h[zf, 1], zero_vals.h[zf, 2]] <- 
    nz_min.h
}

zero_vals.t <- which(current_landscape[,,landscape_predictors[2]] == 0, arr.ind = T)
nz_min.t    <- min(current_landscape[,,landscape_predictors[2]][current_landscape[,,landscape_predictors[2]] != 0])
for (zf in 1:nrow(zero_vals.t)) {
  current_landscape[,,landscape_predictors[2]][zero_vals.t[zf, 1], zero_vals.t[zf, 2]] <- 
    nz_min.t
}

## old names to new names
landscape[[landscape_counter]][,,"urban"]  <- current_landscape[,,landscape_predictors[1]]
landscape[[landscape_counter]][,,"forest"] <- current_landscape[,,landscape_predictors[2]]

```


### LouLand
https://github.com/gaelso/eNFI/blob/c56fe8527a12bbe2ab18295e6bde9c397b7f8dec/data-raw/create-louland.R


https://github.com/gaelso/NFI-modules-2021-lesson1-bookdown/blob/65061e9184b5ec37ee62bf1039f1e181bc7da6ad/R/01-read-data.R

```{r}

## --- Sizes -------------------------------------------------------------- 
length <- 90              ## End results length in km
res    <- 90              ## End results resolution in m
ll     <- length * 10^3   ## Conversion km to m
pp     <- round(ll / res) ## End results pixels number
mp     <- 1000            ## Map pixels number
ff     <- round(pp/mp)    ## Aggregation/disaggregation factor
rr     <- res * ff        ## map resolution in m

.mg = TRUE


## --- Forest types parameters --------------------------------------------
ft_param <- tibble(
  lc_id = c(   5,    4,    3,    2),
  lc    = c("EV", "MD", "DD", "WL"),
  w     = c(0.11, 0.23, 0.08, 0.21),
) %>%
  bind_rows(list(lc_id = 1, lc = "NF", w = 1 - sum(.$w))) %>%
  arrange(lc_id) %>%
  dplyr::mutate(
    hex = c("#edf5e1", "#ffcb9a", "#5cdb95", "#379683", "#00743f")
  )



## --- Land cover parameters extension ------------------------------------
lc_param <- ft_param %>% 
  bind_rows(list(lc_id = c(0, 6), lc = c("WA", "MG"), w = c(0, 0), hex = c("#73c2fb", "#012172"))) %>%
  arrange(lc_id) %>%
  { if (!.mg) dplyr::filter(., lc_id != 6) else .}

#scales::show_col(lc_param$hex)




## Make Topo Raster #######################################################

time1 <- Sys.time()

## --- Make base layers ---------------------------------------------------
#if (!is.null(.seed)) set.seed(.seed)
topo_mpd <- NLMR::nlm_mpd(nrow = mp+5,  ncol = mp+5, resolution = rr, roughness = 0.4, rescale = T, verbose = F)
topo_mpd <- raster::crop(topo_mpd, raster::extent(0, ll, 0, ll))
# topo_mpd
plot(topo_mpd)



 ft_ne <- NLMR::nlm_neigh(nrow = mp/10,  ncol = mp/10, resolution = rr*10, 
                     p_neigh = 0.7, p_empty = 0.1,
                     categories = 5, proportions = ft_param$w, neighbourhood = 8)

```
This doesn't work... 


### WorkshopLand
https://github.com/eco4cast/Statistical-Methods-Seminar-Series/blob/cf138fffd47aa53eb1240f2b91777daf24e7fc60/avgar-smith_issa/extra/sim_habitat.R
```{r}
################################################################X
#-------------ESA Statistical Methods Seminar Series------------X
#------Avgar & Smith -- Integrated Step Selection Analysis------X
#-------------------------07 March 2022-------------------------X
################################################################X


# ... landcover ----
# categories are:
#   - grassland (50%)
#   - forest (30%)
#   - wetland (20%)

set.seed(20220105 + 4)
h4 <- nlm_mosaictess(ncol = 300, nrow = 300, 
                     resolution = 1, germs = 400)

germ2lc <- data.frame(value = sort(unique(values(h4)))) %>% 
  dplyr::mutate(landcover = case_when(
    value < 0.5 ~ 1,
    value < 0.8 ~ 2,
    TRUE ~ 3
  ))

lc_vals <- data.frame(value = values(h4)) %>% 
  left_join(germ2lc)

lc <- setValues(h1, lc_vals$landcover)

# Check
table(values(lc))/(300*300)


### landscapetools::show_landscape(lc)

```



### StackLand 
https://github.com/jmsigner/movement_workshop_public/blob/3e7abd8e157077eedf81fdca8f5e2e62e4440fc1/09%20Multiple%20Animals/09b_walkthrough.R
https://github.com/jmsigner/movement_workshop_public/blob/d72c77aaf3e8489fcc8ace2e137cb3f11882da62/06%20iSSF/06b_walkthrough.R
```{r}
# Simulations ----

# ... Landscape ----
set.seed(1234301)

# Forest
dim <- 500
forest <- NLMR::nlm_gaussianfield(dim, dim) < 0.5
raster::plot(forest)

# and elevation
ele <- NLMR::nlm_gaussianfield(dim, dim)
raster::plot(ele)

covars <- raster::stack(forest, ele)
names(covars)
names(covars) <- c("forest", "elevation")
```
This doesn't work super great - just layers them.. 



## Scrap work... 

```{r}
## need to load...

library("RandomFieldsUtils")
library("devtools")

# This was a real pain in the arse
#library("RandomFieldsUtils")
#library("devtools")
## install_github("cran/RandomFields")
## install_version("RandomFields", "3.3.14")
## install_github("cran/NLMR")


```



source - https://github.com/tnmcfadden/Predictive-Multilayer-Networks/blob/dabcfc0c7bc5a8dbf906d074e0b483c4e9adb5d4/RCode_1_Landscape_simulation.R
```{r}
#################################################_
##### Create a simulated land cover map #####
#################################################_

# Set the basic raster parameters - 30m cells, using UTM coordinate system
r <- raster(nrows=100, ncols=100, xmn=656000, xmx=659000, ymn=5588000, 
            ymx=5591000, resolution=30, crs = CRS("+init=epsg:5362"))
r <- setValues(r, 0)
summary(r)

# Assign land cover (1 = pasture, 2= cropland, 3 = secondary forest, 4 = primary forest)
set.seed(123)
pasture <- makeClass(r, 70, abs(rnorm(70, mean = 100, sd = 100)), val = 1)
set.seed(123)
crops <- makeClass(pasture, 20, abs(rnorm(20, mean = 20, sd = 10)), val = 2, bgr = c(0,1))
set.seed(123)
secondary <- makeClass(crops, 100, abs(rnorm(100, mean = 120, sd = 100)), val = 3)
set.seed(123)
primary <- makeClass(secondary, 5, abs(rnorm(5, mean = 200, sd = 100)), val = 4, bgr = c(0,1,2,3))
rs <- rmSingle(primary)

plot(rs)
```
